---
title: 'Documentation: number of asylum-seekers 2014 by origin and destination'
output: html_document
---

## Introduction into the Shiny App

The shiny app can be found at:
[https://johnzmith.shinyapps.io/ShinyApp]

The shiny app is straight forward to use.
Klick on a destination or origin country to include or exclude it from the 
Sankey plot. The plot displays the number of asylum-seekers in 2014.
Including several countries into the plot gives a direct impression of the magnitude
of the numbers and a comparison is easily understood.


## Origin of the data

The data was downloaded from UNHCR at 11/09/2015:

```{r}
download.file("http://www.unhcr.org/statistics/2014AsylumTrends.zip", "asylum.zip")
unzip("asylum.zip")
```

## Preprocessing and cleaning the data 

The data is hidden in a excel-file with several sheets. The data was cut into
pieces there to fit a normal page. We need to reverse it and read it in several
steps. The country labels will be shortened.

```{r, message =FALSE}
require(XLConnect)
require(stringr)
wk <- loadWorkbook("2014 Asylum Trends annex tables v3.xls") 
df1 <- readWorksheet(wk, sheet=7, startRow=5, endRow=25, colTypes = "character")[,-1]
df2 <- readWorksheet(wk, sheet=7, startRow=27, endRow=47, colTypes = "character")[,-1]
df3 <- readWorksheet(wk, sheet=7, startRow=49, endRow=69, colTypes = "character")[,-1]
df <- data.frame(df1,df2,df3)
df <- data.frame(sapply(df, function(x){str_replace_all(x,"\\.|,", "" )}))
df <- as.matrix(sapply(df, function(x){str_replace_all(x,"\\*", "3" )}))
class(df)<- "numeric"
df <- data.frame(df)
Origin <- readWorksheet(wk, sheet=7, startRow=5, endRow=25)[,1]
Origin[1]<-"Syria"
Origin[4]<-"Serbia/Kosovo"
Origin[9]<-"Iran"
Origin[11]<-"Russia"
Origin[19]<-"El.Salvador"
rownames(df)<-Origin
Origin
Destination <- colnames(df)
Destination[5]<-"Bosnia/Hg."
Destination[18]<-"Iceland"
Destination[21]<-"Japan"
Destination[23]<-"Liechtenstein"
Destination[25]<-"Luxembourg"
Destination[27]<-"Montenegro"
Destination[28]<-"Netherlands"
Destination[33]<-"Korea"
Destination[35]<-"Serbia"
Destination[41]<-"Macedonia"
Destination[42]<-"Turkey"
Destination[44]<-"United.States"
colnames(df) <- Destination
Destination
df[1:6,1:6]
```

The data frame will be put in tidy form, ordered by origin and destination:

```{r, message = FALSE, warning=FALSE}
require(dplyr)
require(tidyr)
dg <- data.frame(Origin=Origin,df) %>% gather(Destination, Number, -Origin) %>%
        arrange(desc(Number)) %>% filter(Number > 0)
head(dg, 10)
```

Both versions were written down into single files:

```{r, eval = FALSE}
write.csv(data.frame(Origin=Origin, df),"asylumtable.csv", row.names = FALSE)
write.csv(dg, "asylum.csv", row.names = FALSE)
```
